Visual Interrogation of Attention-Based Models: \\ Applications in Language Inference and Machine Comprehension

### 2018-05-30 EMNLP-demo-paper/result.tex
update related works
shorten implementation

%\begin{figure*}[t]
%\centering
%\vspace{-2mm}
% \includegraphics[width=1.0\linewidth]{NLI_MC_interface}
%  \vspace{-6mm}
% \caption{
%Illustration of different configurations for the natural language inference and machine comprehension tasks.
%}
%\label{fig:pipelineUpdate}
%\end{figure*}
\section{Applications}
We demonstrate the proposed visualization system on the decomposable attention network~\cite{parikh2016emnlp}
for the NLI task and the bi-directional attention flow network~\cite{Seo2016} for the MC task.

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{NLIexample}
  \vspace{-6mm}
 \caption{
An illustrate of the attention editing process. 
By removing the alignment mistake between the two "green", the originally wrong prediction \emph{entailment} is corrected to \emph{netural}.
}
\label{fig:NLIexample}
\end{figure*}

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predict the relationship between a premise sentence (\textbf{P}) and a hypothesis sentence (\textbf{H}).
%as either \emph{entailment}, \emph{contradiction}, or \emph{neutral}.
%The model can be formalized as the following:
%%\begin{align}
%	$P^\prime, H^\prime = f(P), f(H)$,
%	$\overleftarrow{A} = P^\prime \cdot H^\prime$,
%	$\overrightarrow{A} = H^\prime \cdot P^\prime$,
%	$y = g(P, H, P^\prime, H^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $P$ and $H$ are input embedding matrices for the premise and the hypothesis, $\overleftarrow{A}$
%and $\overrightarrow{A}$ are attentions, and $y$ is the predicted probabilities for candidate classes.
%Our system visualized the bidirectional attentions and their interaction with output distribution over labels.
The model produces two attention matrices, representing alignment from premise to hypothesis, and from hypothesis to premise.

Here we show an example, in which the wrong prediction can be fixed by correcting the attention value.  
%In the proposed system, we overlay the sentence dependency tree with the attention, which enables researchers to conduct comparisons between attention and grammar structure.
%
As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pairs (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is predicted to be entailment, which is incorrect.
By examining the attention, we can see the word \textbf{green} in ``\textbf{green} jacket" is aligned to the \textbf{green} in ``\textbf{green} tea", yet these two ``green'' decorate different nouns. Such a relationship can be confirmed by examining the dependency tree, where  \textbf{green} in \textbf{P} is attached to ``jacket'', whereas \textbf{green} in \textbf{H} is attached to ``tea''. However, the model does not aware of these the grammar information. As a result, the model mistakenly believes the two \textbf{greens} are used to describe the same thing (thus predict \emph{entailment}). 
%
To correct the mistake, we edit the attention value and remove the align between these two "green" (see Figure~\ref{fig:NLIexample}(b)).
As expected, the prediction label is corrected (\emph{neutral}) (see in Figure~\ref{fig:NLIexample}(c)).

>>>>

%\begin{figure*}[t]
%\centering
%\vspace{-2mm}
% \includegraphics[width=1.0\linewidth]{NLI_MC_interface}
%  \vspace{-6mm}
% \caption{
%Illustration of different configurations for the natural language inference and machine comprehension tasks.
%}
%\label{fig:pipelineUpdate}
%\end{figure*}
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{NLIexample}
  \vspace{-6mm}
 \caption{
An illustration of the attention editing process. 
By removing the ``wrong'' alignment between the two ``green", the originally prediction \emph{entailment} is corrected to \emph{netural}.
}
  \vspace{-3mm}
\label{fig:NLIexample}
\end{figure*}

\section{Applications}
We demonstrate the proposed visualization system on the decomposable attention network~\cite{parikh2016emnlp}
for the NLI task and the bi-directional attention flow network~\cite{Seo2016} for the MC task.

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the relationship between a premise sentence (\textbf{P}) and a hypothesis sentence (\textbf{H}).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be correct by editing the attention value.

%The model produces two attention matrices, representing alignment from premise to hypothesis, and from hypothesis to premise.
%as either \emph{entailment}, \emph{contradiction}, or \emph{neutral}.
%The model can be formalized as the following:
%%\begin{align}
%	$P^\prime, H^\prime = f(P), f(H)$,
%	$\overleftarrow{A} = P^\prime \cdot H^\prime$,
%	$\overrightarrow{A} = H^\prime \cdot P^\prime$,
%	$y = g(P, H, P^\prime, H^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $P$ and $H$ are input embedding matrices for the premise and the hypothesis, $\overleftarrow{A}$
%and $\overrightarrow{A}$ are attentions, and $y$ is the predicted probabilities for candidate classes.
%Our system visualized the bidirectional attentions and their interaction with output distribution over labels.

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pairs (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is predicted to be entailment, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two ``green'' decorate different nouns, which lead to the wrong prediction. The grammar structure is captured by the dependency tree, i.e., \emph{green} in \textbf{P} is attached to ``jacket''. However, the model does not has access to the grammar information and mistakenly assume the two \emph{greens} are describing the same thing (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "green" (see Figure~\ref{fig:NLIexample}(b)).
As expected, the prediction label is corrected (\emph{neutral}) (see in Figure~\ref{fig:NLIexample}(c)).

### 2018-05-30 EMNLP-demo-paper/system.tex
update related works
update figures
shorten implementation

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipeline}
 \caption{
 Perturbation-driven interrogation of the end-to-end models that follows the encoder, attention, and classifier structure. The user can generate small perturbation of the input (i.e., replacing synonyms, paraphrasing)
 }
\label{fig:modelPipeline}
\end{figure}

\section{Visualization System}
As illustrated in Figure~\ref{fig:modelPipeline}, many recent end-to-end NLP model follows a similar encoder, attention, and classifier structure, in which the attention stage provide a window for peeking into the model decision making process. However, the static attention alone likely does not tell the whole story, instead we can learn more about the model by examining the dynamic between the input, the attention, and the output. To allow the user to interrogate the relationship between different components of the pipeline, the proposed system utilizes a perturbation-driven exploration strategy, where the user can manually or automatically perturb the input sentence and then view the aggregated prediction information. Beside the perturbation of the input, the system also allows interactive modification of the attention values, where the prediction will instantaneously updated to reflect the changes. 
%
The system consistence several visualization components that can be combined to generate an functional interactive system. In the following section, we will cover some of the key components.

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{sentence}
 \caption{
The interface for showing input sentences. The user can manually alter the words or apply automatic perturbation/paraphrasing of the inputs. In the ``perturbed'' dropdown menu, the blue words are the only not exist in the original sentence.
 }
 \vspace{-3mm}
\label{fig:sentence}
\end{figure}

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{attentionPanels}
  \vspace{-3mm}
 \caption{
Attention visualization. In the graph attention view (a), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. In the matrix attention view (b), the entries of $i^{th}$ row represent the probabilities of words in hypotheses align to the $i^{th}$ word in the premise.
The user can alter the attention values via the pop-up interface illustrated in (c).
We overlay the dependency tree ($a_1$) grammar structure to highlight important words and allow simplification of complex sentence based on the dependency tree.
%
For highly asymmetric attention relationship, we utilized a zoomable hierarchical visual representation (d).
}
\vspace{-3mm}
\label{fig:attentionVis}
\end{figure*}

@@

\subsection{Implementation}
The initial setup cost and learning curve of the tool are often the barriers for user adaptation. To address these challenge, we design the visualization system as a Python library with modularity and easy accessibility in mind.
Instead of using the visualization system as a monolithic standalone application, just like a plotting library (i.e., matplotlib), the different pieces of the visualization (i.e., matrix based attention encoding) can be accessed individually.
% 
Yet, the components can also be combined in any configuration desired by users via a simple API to better fit into one's workflow.
More importantly, the library-based design allows easy integration with the existing model implemented in Python.
%
To create a visualization, users only need to import the library, create an instance of the visualization object, and specify a set of callback functions, such as generating a prediction, accessing attention, to link the visualization to their NLP models. The the code required to create an interactive exploration environment for machine comprehension model is illustrated bellow.

>>>>

\section{Visualization System}
As illustrated in Figure~\ref{fig:modelPipeline}, many recent end-to-end NLP model follows a similar encoder, attention, and classifier structure, in which the attention stage provide a window for peeking into the model decision making process. 
%However, the static attention alone likely does not tell the whole story, instead we can learn more about the model by examining the dynamic between the input, the attention, and the output. 
To allow the user to interrogate the relationship between different components of the pipeline, the proposed system utilizes a perturbation-driven exploration strategy, where the user can manually or automatically perturb the input sentence and then view the aggregated prediction information. Beside the perturbation of the input, the system also allows interactive modification of the attention values, where the prediction will instantaneously updated to reflect the changes. 
%
The system consistence several visualization components that can be combined to generate an functional interactive system. In the following section, we will cover some of the essential components (the usage of other components
%such as a prediction summary statistics the entire development set, as well as a scheme to update the model based 
can be found in the accompany video).

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{sentence}
 \vspace{-2mm}
 \caption{
The interface for showing input sentences. The user can manually alter the words or apply automatic perturbation/paraphrasing of the inputs. In the ``perturbed'' dropdown menu, the blue words are the only not exist in the original sentence.
 }
 \vspace{-1mm}
\label{fig:sentence}
\end{figure}

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{attentionPanels}
  \vspace{-3mm}
 \caption{
Attention visualization. In the graph attention view (a), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. The same values can also be directly visualized in the matrix form (b). 
The user can edit the attention values via the pop-up interface illustrated in (c).
We overlay the dependency tree ($a_1$) grammar structure to highlight important words and allow simplification of complex sentence based on the dependency tree.
%
For highly asymmetric attention relationship, we utilized a zoomable hierarchical visual representation (d). User can focus on individual sentence by clicking the rectangular bars.
}
\vspace{-3mm}
\label{fig:attentionVis}
\end{figure*}

@@

\subsection{Implementation}
The initial setup cost and learning curve of the tool are often the barriers for user adaptation. 
Therefore, instead of design visualization system as a monolithic standalone application, 
we implement the proposed system as a python library with modularity and ease of accessibility in mind.
To fit one's workflow, the different pieces of the visualization (i.e., matrix based attention encoding) can be accessed individually, 
or combined in any configuration desired by users via a simple API.
More importantly, the library-based design allows easy integration with the existing model implemented in python.
%To create a visualization, users only need to import the library, create an instance of the visualization object, and specify a set of callback functions, such as generating a prediction, accessing attention, to link the visualization to their NLP models. 
The the code required to create an interactive exploration environment for machine comprehension model is illustrated bellow.
